\scuba: iterative map-making with the Sub-Millimetre User Reduction Facility

### 2012-09-10 smurf_paper.tex
numerous minor edits to entire paper except for the conclusions

\begin{abstract}
  The Submillimetre Common User Bolometer Array 2 (\scuba) is an
  instrument operating on the 15-m James Clerk Maxwell Telescope,
  consisting of about 5000 bolometers in each of two simultaneous
  imaging bands centred over 450 and 850\,\micron. The camera is
  operated by scanning across the sky and records data at a rate of
  200\,Hz. The data rate is much greater than that of previously
  existing submillimetre cameras, and represents a significant
  analysis challenge. We describe the production of \scuba\ maps using
  the Sub-Millimetre User Reduction Facility (SMURF) in which we have
  adopted a fast, iterative approach to map-making that enables data
  reduction on single, modern, high-end desktop computers, with
  execution times that are shorter than the observing times.  SMURF is
  used in an automated setting, both at the telescope for real-time
  feedback to observers, as well as for the production of generic
  science products at the Canadian Astronomy Data Centre. A suite of
  tools is also provided for interactive data analysis.  Three
  detailed case studies are used to: (i) explore convergence
  properties of the map-maker using simple prior constraints (Uranus
  -- a point source); (ii) achieve the white-noise limit of the
  instrument for faint point-source studies (blank-field survey of the
  Lockman Hole); and (iii) demonstrate that our strategy is capable of
  recovering angular scales up to the size of the array footprint
  (approximately 5\,arcmin) for bright extended sources (star-forming
  region M17).
\end{abstract}

@@

Over the last twenty years, observations in the submillimetre band
(defined here to be 200--1200\,\micron) have helped revolutionise
several important areas of astrophysics, including: the
characterisation of the early stages of star-formation by identifying
the dense, cold regions in molecular clouds where stars may eventually
form; discovering through blind surveys a class of massive dusty
star-forming galaxies in the early ($z>2$) Universe, now referred to
as submillimetre galaxies, or SMGs; and identifying debris disks
around nearby stars, helping us understand the early stages of planet
formation.  With $\sim10,000$ total detectors, \scuba\ is presently the
largest submillimetre camera in the world, and the fastest wide-area
ground-based submillimetre imager.

Submillimetre imaging cameras generally use bolometers, rather than
coherent detectors, to maximise sensitivity. The sensitivity of
bolometers is limited by white photon and phonon noise from the
instrument and ambient backgrounds. The low-frequency noise, however,
is typically dominated by sources which produce slow variations in the
background (e.g., thermal variations within the cryostat, and the
atmosphere for ground-based cameras), and drifts in the readout
electronics. Such noise has a power spectrum $\propto 1/f^\alpha$
($\alpha>0$), and the frequency at which it is comparable to the white
noise level is called the ``$1/f$ knee''. Since the low-frequency
noise is largely correlated between all, or subsets, of the bolometers
in time, it can be suppressed during map-making, since astronomical
signals have the distinct property that they are fixed in a sky
reference frame (assuming they are not time-varying). If successful,
the noise in the resulting map is said to be ``white noise limited'',
meaning that the map noise is uncorrelated spatially, and has an
amplitude that scales as the $\mathrm{NEFD}/\sqrt{t}$, where the NEFD
is the noise-equivalent flux-density (the white noise level of a
bolometer in 1\,s of integration), and $t$ is the amount of
integration time in a map pixel.

@@

There are three general styles of map-making that are relevant to
reducing bolometer data in the literature. The simplest ``direct
methods'' involve some basic processing of the data to remove as much
noise contamination as possible, (e.g., using baseline removal and
other simple filters), and then re-gridding these cleaned data into a
map. Such was the basic recipe for the reduction of chopped data from
SCUBA-2's predecessor SCUBA
\citep{1998ASPC..145..216J,2000ASPC..216..559J}, and MAMBO
\citep[e.g.,][]{omont2001}, another camera from the same
generation. Generally speaking, such methods are fast, although
depending on the science goals and noise properties of the data, they
may not achieve the best noise performance on the angular scales of
interest. In more recent years a method that has been popular for
reducing data in fields of faint point sources with BOLOCAM
\citep[e.g.,][]{laurent2005}, and its younger sibling AzTEC
\citep[e.g.,][]{scott2008}, is ``principal component analysis'' (PCA)
cleaning. A statistical ``black-box'' removes the most correlated
components of the bolometer signals, enabling the detection of
point-sources very close to the theoretical white-noise limits of the
detectors, with reasonable computation times when small numbers of
bolometers are involved (hundreds rather than thousands of
detectors). However, PCA cleaning is not a good generic solution for
producing maps of extended structures, since such sources produce
correlated signals amongst many detectors, and are removed by this
procedure.

The best existing map-making strategies for recovering information on
all angular scales are maximum likelihood techniques, in which the
time-series data are expressed as a sampling of the ``true'' map of
the sky plus noise, and then the equation is inverted to estimate the
map as some weighted linear combination of the data that minimises the
variance.  The first good description of this method appears in
\citet{janssen1992} for the production of maps from the COsmic
Background Explorer (COBE -- the description is relevant despite the
fact that it used a differential radiometer instead of
bolometers). Other descriptions in the CMB literature include
\citet{tegmark1997} and \citet{stompor2002}, and an application to
data from SCUBA is described in \citet{borys2004}.  The downside to
such methods is that they can be both computationally expensive and
have large memory requirements. While for some experimental designs
fast iterative methods for the inversion do exist with minimal memory
usage \citep[e.g.,][]{wright1996}, for the more general map-making
problem involving many detectors, things are significantly complicated
both by the need to measure the cross power-spectra for all unique
pairs of bolometers, as well as performing the inversion itself. The
most promising maximum-likelihood method that may one day be applied
to \scuba\ data is ``SANEPIC'', which successfully reduced maps from
the Balloon-borne Large-Aperture Submillimeter Telescope, involving
hundreds of bolometers, while correctly incorporating inter-bolometer
noise correlations \citep[][]{patanchon2008}.

%  Styles of map-making: least
%squares \citep{janssen1992} maximum likelihood
%\citep[e.g.,][]{patanchon2008}; iterative -- see \citet{johnstone2000}
%implementation of \citet{wright1996} to remove chop,
%\citet{kovacs2008} for SHARC-2 etc.; de-correlation using PCA analysis
%\citep[e.g.,][]{laurent2005,scott2008,aguirre2010}.

The third approach adopted here for \scuba\ is a compromise between
the previous two methods. Under the assumption that a significant
portion of the (predominantly low-frequency) non-white noise sources
can be modelled, an iterative solution is obtained for both the
astronomical image and the parameters of the noise model. Since the
remaining (non-modelled) noise sources are assumed to be white, a
single scalar noise may be calculated for all of the data points from
a given bolometer (since the noise at any instant in time is
uncorrelated with others, and with the data from other bolometers),
greatly simplifying the measurement of noise properties and the
inversion step that is so complicated in the maximum-likelihood
methods. In principle, it should also be possible to recover a wider
dynamic range in angular scales than possible for the simplest direct
methods, with an increase in calculation time that is merely linear in
the number of iterations needed, while using a comparable amount of
memory. A fast iterative method for handling chopped data from CMB
experiments was described in \citet{wright1996}, and later applied to
chopped SCUBA data by \citet{johnstone2000}, although in each case the
chopping renders the data essentially white without the need to model
additional noise components. More similar to the algorithm described
here was the iterative pipeline recipe for fitting and removing
baseline drifts in SCUBA scan-map data as the astronomical image
estimate improved \citep{1999ASPC..172..171J}. The two closest modern
relatives of SMURF are the Comprehensive Reduction Utility for SHARC-2
\citep[CRUSH,][]{kovacs2008}, and the pipeline developed for the
Bolocam Galactic Plane Survey \citep{aguirre2010}.

@@

This paper is organised as follows. We first describe the properties
of \scuba\ data in Section~\ref{sec:data} to motivate the data model
that we have adopted. Next, the details of the SMURF algorithm are
given in Section~\ref{sec:algorithm}. The paper is concluded in
Section~\ref{sec:examples} with three detailed test cases that span
the majority of observation types likely to be undertaken with \scuba,
with an emphasis on the mitigation of divergence problems, and
characterising the output maps: (i) Uranus, a bright, compact source
(Section~\ref{sec:point}); (ii) the Lockman Hole, a blind survey of
faint point-like sources (Section~\ref{sec:cosmo}); and (iii) the
star-forming region M17, including bright, extended emission
(Section~\ref{sec:extended}). All of the data analysed in this paper
are publicly available through the CADC \scuba\ raw-data queries page
for the dates and observation numbers given in the
text\footnote{\url{http://www.cadc-ccda.hia-iha.nrc-cnrc.gc.ca/jcmt/search/scuba2}}.
All of the analysis was performed using the Starlink \textsc{kapuahi}
release from 2012.

%\scuba\ is presently unique in terms of the shear volume of data that
%it produces on a nightly basis, and the fact that it is observing a
%wide variety of science targets for multiple users (unlike dedicated
%CMB experiments, for example). For this reason, SMURF has been
%designed to use a faster, and less memory-intensive algorithm than
%CMB-style maximum likelihood map-makers, although it is less accurate
%at reproducing large-scale structures. The basic procedure is to
%iteratively model and remove most of the bolometer-bolometer
%correlated signal components and independent low-frequency noise, and
%then regrid the residual signals.

% Historically, both radio
%and submillimetre telescope systems have used ``chopping'' to perform
%rapid differences between target sky locations and some reference
%(e.g., another point on the sky, internal calibrators, and different
%frequencies in the case of heterodyne systems). In fact, spatial
%chopping was used by SCUBA, the predecessor to \scuba, to remove
%low-frequency drifts \citep{holland1999}. However, chopping (in
%particular spatial chopping) has the unfortunate side-effect of
%filtering out information on scales larger than the chop-throw
%\textbf{(references)}.For this reason, modern cameras like \scuba\ aim
%to provide a high-degree of inherent detector stability, and rapidly
%slew the telescope so that astronomical signals up to large scales are
%captured in a signal band at considerably higher frequency than the
%residual low-frequency drift.

%In the case of chopped data, ideally the resulting data are completely
%free of low-frequency noise, and therefore the data samples are
%statistically uncorrelated.

%often used various
%``chopping'' methods to remove low-frequency drifts -- rapid
%differences are made between the sky signal at a given position and an
%internal reference signal; the sky signal in two different bands for
%heterodyne systems (frequency chopping); and a different position in
%the sky \citep[a method employed by SCUBA, the predecessor of
%\scuba][]{holland1999}.

@@

Incoming light passes through a beam-splitter, and then bandpass
filters, providing simultaneous illumination and wavelength definition
of both the 450 and 850\,\micron\ focal planes. Each focal plan is
populated by four ``subarrays'' (labeled s4a--s4d at 450\,\micron, and
s8a--s8d at 850\,\micron), each consisting of 32 columns and 40 rows
of bolometers. The bolometers are thermal absorbers coupled to
superconducting transition-edge sensors (TESs) for thermometry;
temperature variations in the TESs produce changing currents, and
therefore varying magnetic fields.  The changing magnetic fields are
detected and amplified using chains of superconducting quantum
interference devices (SQUIDs), and the larger output currents are
digitised. Each detector has its own SQUID for the first-stage of the
amplification, but the remaining stages occur within a common chain of
SQUIDs for each of the 32 columns. All 40 rows are read out in
sequence at a row-visit frequency of about 12\,kHz. Such a high sample
rate is unnecessary to produce maps, so the data are re-sampled to
approximately 200\,Hz before writing to disk.  It should also be noted
that there is an additional 41st row of SQUID readouts that are not
connected to TESs. These ``dark SQUIDs'' track non-thermal noise
sources that are common to each column's amplifier chain. The
relationship between the output digitized current, $I$, and the input
power, $P$, or $dI/dP$ is established using flatfield observations
immediately prior to science observations, in which the output signal
is measured throughout a ramp of the pixel heaters (which provide a
known input power); see Section~2.1 in \citet{dempsey2012} and
Section~5.3 in \citet{holland2012} for more details. Finally, the
conversion to astronomical flux units from power involves a correction
for atmospheric extinction \citep[primarily using the JCMT Water
Vapour Monitor to track line-of-sight opacity variations, see
Section~3 in][]{dempsey2012}, and the application of a flux conversion
factor (FCF) which is established from regular measurements of
calibrators such as Uranus \citep[Section~5 in][]{dempsey2012}.

The predominant noise features in \scuba\ bolometers are: (i) the
effectively white thermal noise produced by the phonon and optical
loading on the absorbers (largely independent from bolometer to
bolometer); and (ii) longer timescale drifts resulting from slow
variations in the mean load (e.g., fridge, telescope, or atmospheric
changes), and also the readout electronics (often correlated amongst
many or all of the bolometers). As such, the power spectra of
bolometer time-series exhibit a ``$1/f^\alpha$'' ($\alpha>0$) shape
below some knee frequency, $f_\mathrm{knee}$, and they are then
relatively flat at higher frequencies. The spectra also include line
features which are mainly produced by aliased high-frequency noise
sources detected during the multiplexed readout stage, as described
above.  Finally, an anti-aliasing filter produces roll-off in the
spectrum at frequencies approaching Nyquist prior to the 200\,Hz
re-sampling.

@@

In order to reduce the impact of noise on the final maps, it is
essential that an appropriate scan strategy be used. First, it is
clearly advantageous that the astronomical signals appear in the
lowest-noise region of the power spectrum, namely frequencies above
$f_\mathrm{knee}$. This is achieved by scanning the telescope at large
slew rates, which are practically limited to about
600\,arcsec\,s$^{-1}$ for the JCMT in order to have reasonable
turn-around overheads (the amount of time spent changing direction and
speed, over the largest elevation range possible). At this maximum
speed there is a sample every 3.0\,arcsec, or approximately one third
of the 450\,\micron\ diffraction-limited full-width at half-maximum
(FWHM) -- a typical rule-of-thumb for adequately sampling a Gaussian
point spread function (PSF). The scan patterns are designed,
primarily, to mitigate the effects of low-frequency $1/f$ drift. The
simplest scan one might use is a raster ``boustrophedon'' pattern,
i.e., a series of parallel passes across the field. However, such a
scan yields extremely poor performance for structures that are
transverse to the scan direction, since a long time will have elapsed
between each parallel step. In addition, other periodic noise features
may conspire with such a regularly-spaced pattern to produce periodic
structures in the map, or ``scan-synchronous'' noise. Scan strategies
that visit each region of the map on as many different time scales and
scan angles as possible, offer the best protection against large-scale
noise; such strategies are said to offer good cross-linking. As a
trade-off between truly random scan patterns and feasibility of
implementation, taking into consideration both mechanical and
scheduling constraints, \scuba\ presently has two primary scan modes:
PONG, which covers rectangular areas larger than the array footprint
in a closed pattern that ``bounces'' off the edges at 45\,degree
angles (and also usually combined with a rotation through a number of
fixed position angles to create a ``rotating PONG'' with even better
cross-linking); and a ``constant velocity daisy'' (CV Daisy) for small
areas (of order the array footprint), in which the telescope moves in
a circle, whose centre also slowly traces out a circle. For a more
complete description of the \scuba\ observing modes, see Section~5 in
\citet{holland2012} and also \citet{2010SPIE.7740E..66K}.

@@

In Fig.~\ref{fig:bolos_mix} we show sample time-series from single
bolometers in each of the 450 and 850\,\micron\ focal planes, as well
as variations in the mixing chamber temperature (as a proxy for the
focal-plane temperature), and the telescope pointing, for two data
sets, before and after the upgrades that followed S2SRO.

@@

Data from the fully commissioned instrument, observation 38 on
20111112, are shown in the right panel of
Fig.~\ref{fig:bolos_mix}. Having solved the fridge oscillation
problem, these data no longer exhibit a correlation with the mixing
chamber (note that variations are still seen in the mixing chamber;
however, such variations do not necessarily reflect changes in the
focal plane temperature). There is minimal correlated low-frequency
noise in these 450\,\micron\ data (note that there is almost no
difference once the common-mode has been subtracted). The
850\,\micron\ channel, however, exhibits a more significant signal
that is obviously correlated with the telescope motion. It is also
correlated amongst many of the detectors, and common-mode removal
corrects it to some extent. Part of the reason that this is seen here,
and not in the S2SRO data set shown, is that the amplitude of the scan
pattern is larger. Most of this ``scan-synchronous'' noise is
attributed to magnetic field pickup, as described in
Section~\ref{sec:magpickup}.

@@

The dotted coloured lines in Fig.~\ref{fig:pspec} show the PSDs for
raw, though flat-fielded and step-corrected (Section~\ref{sec:steps})
data. At each wavelength, and in both data sets, there are clear $1/f$
knees at frequencies ranging from roughly 1 to 2\,Hz, followed by a
predominantly white spectrum punctuated by line features at higher
frequencies, and finally a roll-off caused by the anti-aliasing filter
above $\gsim 70$\,Hz. As indicated in the previous section, the
correlation between the low-frequency components of the different
bolometer signals is large. The solid black lines in
Fig.~\ref{fig:pspec} indicate the PSDs of the common-modes
$\mathbf{c}(t)$ at each wavelength, which reproduce much of the
low-frequency structure, as well as some of the higher-frequency line
features. The $\mathbf{c}(t)$ otherwise drop substantially below the
individual bolometer PSDs at high-frequency, as expected if the
bolometers are dominated by uncorrelated white noise in that part of
the spectrum. The common-mode signals are fit to each bolometer time
series and removed as in Fig.~\ref{fig:bolos_mix}, and the resulting
PSDs are shown as solid coloured lines.  For reference, the top
horizontal axes have been converted to angular scale assuming the scan
speeds of each observation (120\,arcsec\,s$^{-1}$ in the left-hand
plot, and 190\,arcsec\,s$^{-1}$ in the right-hand plot). The power
spectra of unresolved point-sources are also shown as dashed lines in
each band (arbitrarily normalised), showing that the smallest features
resolvable by the telescope are only minimally affected by the excess
noise in the line features at these scan speeds (this may not be the
case at higher scan speeds).

@@

Raw data from the fully-commissioned instrument (right-hand panel of
Fig.~\ref{fig:pspec}) generally have less significant $1/f$ noise as
compared with the S2SRO data. The residual low-frequency noise is
considerably less correlated amongst detectors (larger spread in the
dotted coloured lines, particularly noticeable in these 450\,\micron\
data), leading to a less drastic improvement upon common-mode
removal. However, since the data are less dominated by low-frequency
noise to begin with, the signals generally require less low-frequency
filtering to produce maps, and therefore retain larger-scale
structures than with the S2SRO data. Furthermore, there are generally
fewer line features in the PSDs of bolometers in the
fully-commissioned instrument. Finally, these two data sets illustrate
that the white noise performance (NEFD) is in fact similar before and
after the upgrades for these two subarrays (s4a at 450\,micron, and
s8b at 850\,\micron). The main improvements are reduction in
correlated and line noise features mentioned above, and the larger
number of working bolometers (approximately a factor of 4).

@@

@@

A method that has been used to remove correlated noise as part of the
map-making procedure for BOLOCAM and AzTEC is Principal Component
Analysis \citep[PCA,][]{laurent2005,scott2008,perera2008}. Here we use
PCA to explore correlated signals in \scuba\ data.

@@

\begin{itemize}

\end{itemize}

@@

\item smooth the gradient values to determine the local mean gradient
and subtract this local mean from the total gradient to get residual
gradient;

@@

\end{itemize}

>>>>

\begin{abstract}
  The Submillimetre Common User Bolometer Array 2 (\scuba) is an
  instrument operating on the 15-m James Clerk Maxwell Telescope,
  consisting of about 5000 bolometers in each of two simultaneous
  imaging bands centred over 450 and 850\,\micron. The camera is
  operated by scanning across the sky and records data at a rate of
  200\,Hz. This rate is much greater than that of previously existing
  submillimetre cameras, and represents a significant analysis
  challenge. We describe the production of \scuba\ maps using the
  Sub-Millimetre User Reduction Facility (SMURF) in which we have
  adopted a fast, iterative approach to map-making that enables data
  reduction on single, modern, high-end desktop computers, with
  execution times that are shorter than the observing times.  SMURF is
  used in an automated setting, both at the telescope for real-time
  feedback to observers, as well as for the production of generic
  science products at the Canadian Astronomy Data Centre. Three
  detailed case studies are used to: (i) explore convergence
  properties of the map-maker using simple prior constraints (Uranus
  -- a point source); (ii) achieve the white-noise limit for faint
  point-source studies (extragalactic blank-field survey of the
  Lockman Hole); and (iii) demonstrate that our strategy is capable of
  recovering angular scales comparable to the size of the array
  footprint (approximately 5\,arcmin) for bright extended sources
  (star-forming region M17).
\end{abstract}

@@

Over the last twenty years, observations in the submillimetre band
(defined here to be 200--1200\,\micron) have helped revolutionise
several important areas of astrophysics, including: discovering
through blind surveys a class of massive dusty star-forming galaxies
in the early ($z>2$) Universe, now referred to as submillimetre
galaxies, or SMGs; the characterisation of the early stages of
star-formation by identifying the dense, cold regions in molecular
clouds where stars may eventually form; and identifying debris disks
around nearby stars, helping us understand the early stages of planet
formation.  With $\sim10,000$ total detectors, \scuba\ is presently
the largest submillimetre camera in the world, and the fastest
wide-area ground-based submillimetre imager.

Submillimetre imaging cameras generally use bolometers, rather than
coherent detectors, to maximise sensitivity. The sensitivity of
bolometers is limited by white photon and phonon noise from the
instrument and ambient backgrounds. The low-frequency noise, however,
is typically dominated by sources which produce slow variations in the
background (e.g., thermal variations within the cryostat, and the
atmosphere for ground-based cameras), and drifts in the readout
electronics. Such noise has a power spectrum $\propto 1/f^\alpha$
($\alpha>0$), and the frequency at which it is comparable to the white
noise level is called the ``$1/f$ knee''. Since the low-frequency
noise is largely correlated between all, or subsets, of the bolometers
in time, it can be suppressed during map-making, since astronomical
signals have the distinct property that they are fixed in a sky
reference frame (assuming they are not time-varying). If successful,
the noise in the resulting map is said to be ``white noise limited'',
meaning that it is uncorrelated spatially, and has an amplitude that
scales as the $\mathrm{NEFD}/\sqrt{t}$, where the NEFD is the
noise-equivalent flux-density (the white noise level of a bolometer in
1\,s of integration), and $t$ is the amount of integration time in a
map pixel.

@@

There are three general styles of map-making that are relevant to
reducing bolometer data in the literature. The simplest ``direct
methods'' involve some basic processing of the data to remove as much
noise contamination as possible, (e.g., using baseline removal and
other simple filters), and then re-gridding these cleaned data into a
map. Such was the basic recipe for the reduction of chopped data from
SCUBA-2's predecessor SCUBA
\citep{1998ASPC..145..216J,2000ASPC..216..559J}, and MAMBO
\citep[e.g.,][]{omont2001}, another camera from the same
generation. Generally speaking, such methods are fast, although
depending on the science goals and noise properties of the data, they
may not achieve the best noise performance on the angular scales of
interest. In more recent years a method that has been popular for
reducing data in fields of faint point sources with Bolocam
\citep[e.g.,][]{laurent2005}, and its younger sibling the Aztronomical
Thermal Emission Camera \citep[AzTEC, e.g.,][]{scott2008}, is
``principal component analysis'' (PCA) cleaning. A statistical
``black-box'' removes the most correlated components of the bolometer
signals, enabling the detection of point-sources very close to the
theoretical white-noise limits of the detectors, with reasonable
computation times when small numbers of bolometers are involved
(hundreds rather than thousands of detectors). However, PCA cleaning
is not a good generic solution for producing maps of extended
structures, since such sources produce correlated signals amongst many
detectors, and are removed by this procedure. Furthermore, in the case
of \scuba\, performing PCA on even a single subarray (1280 detectors,
nominally) can be prohibitively slow.

The best existing map-making strategies for recovering information on
all angular scales are maximum likelihood techniques, in which the
time-series data are expressed as a sampling of the ``true'' map of
the sky plus noise, and then the equation is inverted to estimate the
map as some weighted linear combination of the data that minimises the
variance.  The first good description of this method appears in
\citet{janssen1992} for the production of maps from the COsmic
Background Explorer (COBE -- the description is relevant despite the
fact that it used a differential radiometer instead of
bolometers). Other descriptions in the experimental cosmic microwave
background (CMB) literature include \citet{tegmark1997} and
\citet{stompor2002}, and an application to data from SCUBA is
described in \citet{borys2004}.  The downside to such methods is that
they can be both computationally expensive and have large memory
requirements. While for some experimental designs fast iterative
methods for the inversion do exist without requiring excessive amounts
of memory \citep[e.g.,][]{wright1996}, for the more general map-making
problem, involving many detectors, things are significantly
complicated both by the need to measure the cross power spectra for
all unique pairs of bolometers, as well as performing the inversion
itself. The most promising maximum-likelihood method that may one day
be applied to \scuba\ data is ``SANEPIC'', which successfully reduced
maps from the Balloon-borne Large-Aperture Submillimeter Telescope,
involving hundreds of bolometers, while correctly incorporating
inter-bolometer noise correlations \citep[][]{patanchon2008}.

The third approach adopted here for \scuba\ is a compromise between
the previous two methods. Under the assumption that a significant
portion of the (predominantly low-frequency) non-white noise sources
can be modelled, an iterative solution is obtained for both the
astronomical image and the parameters of the noise model. Since the
remaining (non-modelled) noise sources are assumed to be white, a
single scalar \rms\ may be calculated for all of the data points from
a given bolometer to characterise its noise distribution (since the
noise at any instant in time is uncorrelated with others, and with the
data from other bolometers), greatly simplifying the measurement of
noise properties and the inversion step that is so complicated in the
maximum-likelihood methods. In principle, it should also be possible
to recover a wider dynamic range in angular scales than possible for
the simplest direct methods, with an increase in calculation time that
is merely linear in the number of iterations needed, while using a
comparable amount of memory. The iterative method of
\citet{wright1996} specific to chopped (CMB) data was applied to SCUBA
observations by \citet{johnstone2000}, although in this case the
chopping removes most of the low-frequency noise, and greatly
simplifies the problem (at the expensive of large-scale structure in
the data). More similar to the algorithm described here was the
iterative pipeline recipe for fitting and removing baseline drifts in
SCUBA scan-map data as the astronomical image estimate improved
\citep{1999ASPC..172..171J}. The two closest modern relatives of SMURF
are the Comprehensive Reduction Utility for SHARC-2
\citep[CRUSH,][]{kovacs2008}, and the pipeline developed for the
Bolocam Galactic Plane Survey \citep[][which uses iterative PCA
cleaning to avoid the suppression of large-scale
structure]{aguirre2010}.

@@

This paper is organised as follows. We first describe the properties
of \scuba\ data, including a principal component analysis to reveal
correlated noise features, in Section~\ref{sec:data}. Next, the
details of the SMURF algorithm (pre-processing steps and the iterative
solution) are given in Section~\ref{sec:algorithm}. The paper is
concluded in Section~\ref{sec:examples} with three detailed test cases
that span the majority of observation types likely to be undertaken
with \scuba, with an emphasis on the mitigation of divergence
problems, and characterising the output maps: (i) Uranus, a bright,
compact source (Section~\ref{sec:point}); (ii) the Lockman Hole, a
blind survey of faint point-like sources (Section~\ref{sec:cosmo});
and (iii) the star-forming region M17, including bright, extended
emission (Section~\ref{sec:extended}). All of the data analysed in
this paper are publicly available through the CADC \scuba\ raw-data
queries page for the dates and observation numbers given in the
text\footnote{\url{http://www.cadc-ccda.hia-iha.nrc-cnrc.gc.ca/jcmt/search/scuba2}}.
All of the analysis was performed using the Starlink \textsc{kapuahi}
release from 2012.

@@

Incoming light passes through a beam-splitter, and then bandpass
filters, providing simultaneous illumination and wavelength definition
of both the 450 and 850\,\micron\ focal planes. Each focal plan is
populated by four ``subarrays'' (labeled s4a--s4d at 450\,\micron, and
s8a--s8d at 850\,\micron), each consisting of 32 columns and 40 rows
of bolometers. The bolometers are thermal absorbers coupled to
superconducting transition-edge sensors (TESs) for thermometry.
Temperature variations in the TESs produce changing currents, and
therefore varying magnetic fields, which are detected and amplified
using chains of superconducting quantum interference devices (SQUIDs),
before the larger output currents are digitised. Each detector has its
own SQUID for the first-stage of the amplification, but the remaining
stages occur within a common chain of SQUIDs for each of the 32
columns. All 40 rows are read out in sequence at a row-visit frequency
of about 12\,kHz. Such a high sample rate is unnecessary to produce
maps, so the data are re-sampled to approximately 200\,Hz before
writing to disk. This rate provides a sample every 3.0\,arcsec, or
approximately one third of the 450\,\micron\ diffraction-limited
full-width at half-maximum (FWHM) -- a typical rule-of-thumb for
adequately sampling a Gaussian point spread function (PSF). There is
an additional 41st row of SQUID readouts that are not connected to
TESs. These ``dark SQUIDs'' track non-thermal noise sources that are
common to each column's amplifier chain. The relationship between the
output digitized current, $I$, and the input power, $P$, or $dI/dP$ is
established using flatfield observations immediately prior to science
observations, in which the output signal is measured throughout a ramp
of the pixel heaters (which provide a known input power); see
Section~2.1 in \citet{dempsey2012} and Section~5.3 in
\citet{holland2012} for more details. Finally, the conversion to
astronomical flux units from power involves a correction for
atmospheric extinction \citep[primarily using the JCMT Water Vapour
Monitor to track line-of-sight opacity variations, see Section~3
in][]{dempsey2012}, and the application of a flux conversion factor
(FCF) which is established from regular measurements of calibrators
such as Uranus \citep[Section~5 in][]{dempsey2012}.

In addition to uncorrelated white noise, and low-frequency (and often
correlated from bolometer to bolometer) $1/f$ drifts, bolometer power
spectra include a roll-off in their spectra at frequencies approaching
Nyquist, which is due to the anti-aliasing filter that forms part of
the 200\,Hz re-sampling stage. There are also line features which are
thought to be produced at frequencies far above the final 200\,Hz
sample rate, and are aliased to lower frequencies during the
multiplexed readout stage before they can be removed by the
anti-aliasing filter.

@@

In order to reduce the impact of low-frequency noise on the final
maps, \scuba\ scan strategies have been designed to provide: (i) good
cross-linking (visiting every point of the mapped area on several
different time scales, and at different scan angles); and (ii) minimal
accelerations to reduce turn-around overheads (which could be quite
large given the 600\,arcsec\,s$^{-1}$ maximum scan speed). For areas
larger than the array footprint a closed ``PONG'' pattern is used, in
which the boresight travels in approximately straight lines and
``bounces'' off the edges at 45\,degree angles until the area is
uniformly filled in. It is also usually combined with a rotation
through a number of fixed position angles to create a ``rotating
PONG'' with even better cross-linking. For smaller areas (of order the
array footprint, or pont sources), in which the PONG turn-around
overheads would be large, a ``constant velocity daisy'' (CV Daisy) is
used.  Here the telescope moves in a circle, whose centre also slowly
traces out a small circle. For a more complete description of the
\scuba\ observing modes, see Section~5 in \citet{holland2012} and also
\citet{2010SPIE.7740E..66K}.

@@

In Fig.~\ref{fig:bolos_mix} we show sample time-series from single
bolometers in each of the 450 and 850\,\micron\ focal planes, as well
as variations in the mixing chamber temperature (though not located in
the focal plane itself, it is certainly correlated with the
temperature of the detectors), and the telescope pointing, for two
data sets, before and after the upgrades that followed S2SRO.

@@

Data from the fully commissioned instrument, observation 38 on
20111112, are shown in the right panel of
Fig.~\ref{fig:bolos_mix}. Having solved the fridge oscillation
problem, these data no longer exhibit a correlation with the mixing
chamber (however, note that variations are still seen in the mixing
chamber signal; such variations do not necessarily reflect changes in
the focal plane temperature). There is minimal correlated
low-frequency noise in these 450\,\micron\ data (note that there is
almost no difference once the common-mode has been subtracted). The
850\,\micron\ channel, however, exhibits a more significant signal
that is obviously correlated with the telescope motion. It is also
correlated amongst many of the detectors, and common-mode removal
corrects it to some extent. Part of the reason that this is seen here,
and not in the S2SRO data set shown, is that the amplitude of the scan
pattern is larger. Most of this ``scan-synchronous'' noise is
attributed to magnetic field pickup, as described in
Section~\ref{sec:magpickup}.

@@

The dotted coloured lines in Fig.~\ref{fig:pspec} show the PSDs for
raw, though flat-fielded and step-corrected (Section~\ref{sec:steps})
data. At each wavelength, and in both data sets, there are clear $1/f$
knees at frequencies ranging from roughly 1 to 2\,Hz, followed by a
predominantly white spectrum punctuated by line features at higher
frequencies, and finally a roll-off caused by the anti-aliasing filter
above $\gsim 70$\,Hz. The correlation between the low-frequency
components of the different bolometer signals is large. The solid
black lines in Fig.~\ref{fig:pspec} indicate the PSDs of the
common-modes $\mathbf{c}(t)$ at each wavelength, which reproduce much
of the low-frequency structure, as well as some of the
higher-frequency line features. The $\mathbf{c}(t)$ otherwise drop
substantially below the individual bolometer PSDs at high-frequency,
as expected if the bolometers are dominated by uncorrelated white
noise in that part of the spectrum. The common-mode signals are fit to
each bolometer time series and removed as in Fig.~\ref{fig:bolos_mix},
and the resulting PSDs are shown as solid coloured lines.  For
reference, the top horizontal axes have been converted to angular
scale assuming the scan speeds of each observation
(120\,arcsec\,s$^{-1}$ in the left-hand plot, and
190\,arcsec\,s$^{-1}$ in the right-hand plot). The power spectra of
unresolved point-sources are also shown as dashed lines in each band
(arbitrarily normalised), showing that the smallest features
resolvable by the telescope are only minimally affected by the excess
noise in the line features at these scan speeds (this may not be the
case at higher scan speeds).

@@

Raw data from the fully-commissioned instrument (right-hand panel of
Fig.~\ref{fig:pspec}) generally have less significant $1/f$ noise as
compared with the S2SRO data, and it is considerably less correlated
amongst detectors (larger spread in the dotted coloured lines,
particularly noticeable in these 450\,\micron\ data), leading to a
less drastic improvement upon common-mode removal. However, since the
data are less dominated by low-frequency noise to begin with, the
signals generally require less agressive high-pass filtering to
produce maps, and therefore retain larger-scale structures than with
the S2SRO data. Furthermore, there are generally fewer line features
in the PSDs of bolometers in the fully-commissioned
instrument. Finally, these two data sets illustrate that the white
noise performance (NEFD) is in fact similar before and after the
upgrades for these two subarrays (s4a at 450\,micron, and s8b at
850\,\micron). The main improvements are a reduction in correlated and
line noise features mentioned above, the larger number of working
bolometers and field-of-view (approximately a factor of 4).

@@

\begin{figure}
\centering
\includegraphics[width=\linewidth]{magpickup.pdf}
\caption{Evidence for significant magnetic field pickup for
  observation 16 on 20100228 at 450\,\micron\ (s4a subarray).  The top
  panel shows two un-flatfielded (but mean-subtracted and step
  corrected) bolometer time-series from the same column, with a 200
  sample boxcar smooth (approximately 1\,s), illustrating that they
  are dominated by a similar signal with opposite signs. The second
  panel shows the dark squid signal for the column, also
  mean-subtracted and with the same boxcar smooth. The bottom panel
  shows the azimuthal and elevation offsets from the map centre (mean
  azimuth and elevation 171.9$^\circ$ and 68.0$^\circ$,
  respectively). Only the azimuthal signal is obviously correlated
  with the dark squids and bolometer signals, which suggests a
  magnetic field stationary with respect to the telescope dome as the
  source, since its direction with respect to the cryostat only
  changes with azimuthal motion.}
\label{fig:magpickup}
\end{figure}

@@

A method that has been used to remove correlated noise as part of the
map-making procedure for Bolocam and AzTEC is Principal Component
Analysis \citep[PCA,][]{laurent2005,scott2008,perera2008}. Here we use
PCA to explore correlated signals in \scuba\ data.

@@

@@

\item smooth the gradient values to determine the local mean gradient
and subtract this local mean from the total gradient to get the
residual gradient;

@@

\subsubsection{Additional bad bolometer rejection}
\label{sec:flagbadbol}
